---
title: "p8105_hw3_zj2379_2"
author: "Zheshu Jiang"
date: "2023-10-15"
output: github_document
---
problem 1 part 1 for first try, I have done the hw two days before, but I tried to make some changes, i found i could only update rmd file but not md file, so i am doing it again, my previous hw3 link is https://github.com/jzs43/p8105_hw3_zj2379/tree/main for reference

## Problem1
```{r}
library(p8105.datasets)
data("instacart")
library(tidyverse)
library(ggridges)
library(patchwork)
```

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

There are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered.

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```


Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2

```{r}
#load the data
library(p8105.datasets)
data("brfss_smart2010")
brfss_smart2010
```
First, do some data cleaning:
format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”
```{r}
#do some data cleaning following the above guideline
brfss_smart = 
  brfss_smart2010 |> 
  as_tibble()
brfss_smart = 
  brfss_smart |>
  janitor::clean_names() |>
  rename(state = locationabbr, location_state = locationdesc) |>
  filter(topic %in% "Overall Health") |>
  mutate(response = fct_relevel(response, "Poor","Fair","Good","Very good","Excellent"))

```


