---
title: "p8105_hw3_zj2379_2"
author: "Zheshu Jiang"
date: "2023-10-15"
output: github_document
---
problem 1 part 1 for first try, I have done the hw two days before, but I tried to make some changes, i found i could only update rmd file but not md file, so i am doing it again, my previous hw3 link is https://github.com/jzs43/p8105_hw3_zj2379/tree/main for reference

## Problem1
```{r}
library(p8105.datasets)
data("instacart")
library(tidyverse)
library(ggridges)
library(patchwork)
```

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

There are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered.

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```


Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2

```{r}
#load the data
library(p8105.datasets)
data("brfss_smart2010")
brfss_smart2010
```
First, do some data cleaning:
format the data to use appropriate variable names;
focus on the “Overall Health” topic
include only responses from “Excellent” to “Poor”
organize responses as a factor taking levels ordered from “Poor” to “Excellent”
```{r}
#do some data cleaning following the above guideline
brfss_smart = 
  brfss_smart2010 |> 
  as_tibble()
brfss_smart = 
  brfss_smart |>
  janitor::clean_names() |>
  rename(state = locationabbr, location_state = locationdesc) |>
  filter(topic %in% "Overall Health") |>
  mutate(response = fct_relevel(response, "Poor","Fair","Good","Very good","Excellent"))

```

```{r}
# keep only data from 2002 to see which states were observed at 7 or more locations
brfss_smart |>
  filter(year == "2002") |>
  group_by(state) |>                                                    
  summarize(n = n_distinct(location_state)) |>  
  filter(n >= 7)
```
This plot shows a line for each state across years from 2002 to 2010.
In 2002, CT, FL, MA, NC, Nj, PA were observed at 7 or more locations.

```{r}
# keep only data from 2010 to see which states were observed at 7 or more locations
brfss_smart |>
  filter(year == "2010") |>
  group_by(state) |>                                                    
  summarize(n = n_distinct(location_state)) |>  
  filter(n >= 7)
```

In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, HY, OH, PA, SC, TX, WA were observed at 7 or more locations.

make a plot showing a line for each state across years
```{r}
  brfss_smart |>
  filter(response=="Excellent")|>
  select(year, state, data_value)|>
  group_by(year, state)|>
  summarise(average_data_value = mean(data_value))|>
  ggplot(aes(x = year, y = average_data_value, color = state, group = state)) +
  geom_line()+
  labs(
    x = "Year",
    y = "Average Data Value",
    title = "Average Value Over Time Within a State"
  )
ggsave("line plot for each state.pdf")
  
```
This plot shows a line for each state across years from 2002 to 2010.


Make a two-panel plot showing, for the years 2006, and 2010
```{r}
brfss_smart |>
  group_by(year,response,state,data_value) |>
  filter(year %in% c("2006","2010")) |>
  filter(state %in% "NY")|>
  ggplot(aes(x = response, y = data_value)) + 
  geom_boxplot() + 
  facet_grid(. ~ year)
  ggsave("two-panel plot for 2006 and 2010.pdf")
```
This plot shows the distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State in 2006 and 2010 respectively. Based on the graphs, the group with "poor" responses has the lowest data values at both years, good and very good responses have relatively high data values.

## Problem3

```{r}
nhanes_covar = 
  read_csv("nhanes_covar.csv",skip = 4) |>
  janitor::clean_names() |>
  # exclude participants less than 21 years of age
  filter(age > 21) |>
  mutate(
    sex = recode(sex, "1" = "male", "2" = "female"),
      education = recode(
      education, 
      "1" = "Less than high shcool", 
      "2" = "High school equivalent",
      "3" = "More than high school"),
  )|>
 mutate(
   sex = as.factor(sex), 
   education = as.factor(education)
 )|>
  #exclude those observations with missing demographic data
  na.omit() 
nhanes_covar
```
```{r}
nhanes_accel = 
  read_csv("nhanes_accel.csv") |>
  janitor::clean_names()|>
   pivot_longer(                                           
    min1:min1440,
    names_to = "number",
    values_to = "counts",
    names_prefix = "min"
  )
nhanes_accel
```

```{r}
# combine the two datasets
nhanes_df = left_join(nhanes_covar,nhanes_accel,by = "seqn")
```

Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. 
```{r}
#make a table
education_table=select(nhanes_covar, sex | education)  
table(education_table)
```

At the High school equivalent education level, the number of male is 34 and the number of female is 23, more male than female. At less than high school education level,the number of male is 28 and the number of female is 27, more male than female. At more than high school education level,the number of male is 59 and the number of female is 54, more male than female.

```{r}
nhanes_covar |>
ggplot(aes(x = age, fill = sex )) +
  geom_density(alpha = .5) +
  facet_grid(. ~ education)+
  labs( title = "Age Distribution for Sex for three Education levels",
    y = "age",
    x = "density")
ggsave("Age Distribution for Sex for three Education levels.pdf")
```

Based on the graphs, at the high school equivalent education level, the number of females exceeds the number of males as age increases. At the less than high school education level, the number of females first exceeds the number of males then becomes lower than that of males as age increases. At the more than high school education level, the number of females is originally higher than the number of males and then becomes lower than that of males as age increases.Overall, females and younger people have relatively higher education level than other age groups.

create a total activity variable for each participant
```{r}
accel_clean = 
  nhanes_df |>                                                                  
  group_by(seqn, sex, age, education) |>                        
  summarize(
    total_activity = sum(counts)                                
  )
accel_clean|>
ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ education)+
  labs(
    title = "total activities against age",
    x = "age",
    y = "total activities",
  ) 
ggsave("total activities against age.pdf")
```

At high school equivalent and more than high school education level, overall females activity is higher than males for people 22yr+. At less than high school education level, the overall males activity is higher than females between 40 and 80, but lower than that of females between 20 and 40. At more than high school education level,overall females activity is higher than males at all age groups. No matter education level and gender, total activity decreases as the age increases.


A three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex.
```{r}
nhanes_df |>
ggplot(aes(x = number, y = counts, group = seqn, color = sex)) +
  geom_line(alpha = .2) +
  facet_grid(. ~ education) +
  geom_smooth(aes(group = sex), se = FALSE)+
  labs(
    title = "total activities over 24-hour",
    x = "time(mins)",
    y = "total activities",
  ) 
ggsave("total activities over 24-hour.pdf")
```

Based on the plot, the 24-hour activity time courses for high school equivalent and less than high school education levels are similar. At more than high school education level, there are two peaks of activities. In addition, females have higher activities than males do.From the smooth trends, we could see a decrease for every education level. 

